---
title: "Presentation Visualizations"
author: "Dylan Jacobs, Jun Pu, and Ariel Lutati"
date: "2024-03-07"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Jun's plot
# Import packages
```{r message=FALSE, warning=FALSE}
library(gapminder)
library(ggplot2)
library(plotly)
library(dplyr)
library(readxl)
library(leaflet)
library(sf)
library(RColorBrewer)
library(shiny)
```

# Data Cleaning

```{r}
#selected_columns <- c("Country name", "Year", "Adult population", "Region", "Income group", "Saved at a financial institution (% age 15+)")
df_findex <- read_excel("DatabankWide.xlsx")
df_gdp <- read.csv("df_gdpPC.csv")
```

# Select the variabels of interest
```{r}
df_findex_var <- df_findex[c("Country name", "Year", "Region", "Adult populaiton", "Income group", "Made or received a digital payment (% age 15+)", "Financial institution account (% age 15+)")]

df_gdp_var <- df_gdp[c("country", "year", "gdpPercap", "lifeExp")]
```


# Rename variables to make it look nicer
```{r}
# Create a new dataframe df_rename
df_findex_new <- df_findex_var

# Rename "Made or received a digital payment (% age 15+)" to "payment_digital"
colnames(df_findex_new)[colnames(df_findex_new) == "Made or received a digital payment (% age 15+)"] <- "payment_digital"

# Rename "Financial institution account (% age 15+)" to "account_fin"
colnames(df_findex_new)[colnames(df_findex_new) == "Financial institution account (% age 15+)"] <- "account_fin"

# Rename "Country name" to "country"
colnames(df_findex_new)[colnames(df_findex_new) == "Country name"] <- "country"

# Rename "Adult populaiton" to "population"
colnames(df_findex_new)[colnames(df_findex_new) == "Adult populaiton"] <- "population"

# Rename "Income group" to "group_income"
colnames(df_findex_new)[colnames(df_findex_new) == "Income group"] <- "group_income"

# Rename "Year" to "year"
colnames(df_findex_new)[colnames(df_findex_new) == "Year"] <- "year"

# Rename "Region" to "region"
colnames(df_findex_new)[colnames(df_findex_new) == "Region"] <- "region"

# Remove N/A values
df_findex_new <- na.omit(df_findex_new)


# Check variable type
# str(df_rename)
```

# Keep observations from years 2014, 2017, and 2021
```{r}
# Assuming 'year' is the column containing the year information
df_gdp_var_limit <- subset(df_gdp_var, year %in% c(2014, 2017, 2021))
df_findex_limit <- subset(df_findex_new, year %in% c(2014, 2017, 2021))
```



# Merge the two datasets
```{r}
# by 'country' and 'year'
merged_data <- merge(df_findex_limit, df_gdp_var_limit, by = c("year", "country"))
```


# Create a basic scatter plot

The following scatterplot shows the relationship between percentage of population that has made/received a digital payment and percentage of population with an account at a financial institution. From the graph, we can observe that high income countries have a larger percentage with a financial account and a large percentage of involvement in digital payments.
```{r}
plot_dig_account <- merged_data %>%
  plot_ly(x = ~account_fin, y = ~payment_digital,
          size = ~gdpPercap, sizes = c(5, 30), 
          color = ~group_income, text = ~paste(country, "<br>", year),
          mode = "markers", marker = list(sizemode = "diameter", opacity = 0.6)) %>%
  layout(title = "Financial Account vs. Savings at Financial Institution",
         xaxis = list(title = "Percentage with an account at a Financial Institution"),
         yaxis = list(title = "Percentage made or received a digital payment"),
         showlegend = TRUE)

# Show the plot
plot_dig_account

```


```{r message=FALSE, warning=FALSE}
# Create an interactive scatter plot with a slider for 'year'
# Define the UI
ui <- fluidPage(
  sliderInput("year_slider", "Select Year", min = min(merged_data$year), max = max(merged_data$year), value = min(merged_data$year), step = 1),
  plotlyOutput("scatter_plot")
)

# Define the server
server <- function(input, output) {
  filtered_data <- reactive({
    merged_data[merged_data$year == input$year_slider, ]
  })

  output$scatter_plot <- renderPlotly({
    plot_ly(data = filtered_data(),
            x = ~account_fin, y = ~payment_digital,
            color = ~group_income, text = ~paste(country, "<br>", year),
            size = ~gdpPercap, sizes = c(5,20),
            mode = "markers", marker = list(sizemode = "diameter", opacity = 0.6)) %>%
      layout(title = "Financial Account vs. Savings at Financial Institution",
             xaxis = list(title = "Percentage with an account at a Financial Institution"),
             yaxis = list(title = "Percentage made or received a digital payment"),
             showlegend = TRUE)
  })
}

# Run the Shiny app
shinyApp(ui, server)

```



```{r}
# Add a new column with corresponding regions
library(countrycode)
merged_data$better_region <- countrycode(merged_data$country, origin = "country.name", destination = "region")
```

```{r}
new_plot <- plot_ly(merged_data, x = ~better_region, y = ~payment_digital, color = ~country, type = 'bar') %>%
  layout(
    title = "Comparison of Regions by Income Group and Digital Payment",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Value"),
    barmode = 'group'
  )

new_plot
```
```{r}
top_digi_ctrs <- merged_data %>% arrange(better_region, desc(payment_digital)) %>%
  group_by(year, better_region) %>%
  slice_head(n = 5)

top_digi_ctrs %>% plot_ly(x = ~better_region, y = ~payment_digital, color = ~country) %>%
  add_histogram(hoverinfo = "x+y") %>%
  group_by(year, better_region) %>%
  summarise(n = n()) %>%
  layout(
    barmode = 'stack',
    title = "Top 5 Most Digital Countries in Each Region",
  xaxis = list(title = "Country"),
  yaxis = list(title = "Payment Digital"),
  sliders = list(
    list(
      active = 0,
      steps = lapply(unique(top_digi_ctrs$year), function(year) {
        list(
          label = year,
          method = "update",
          args = list(list(visible = top_digi_ctrs$year == year),
                      list(title = paste("Top 5 Most Digital Countries in Each Region - ", year)))
        )
      })
    )
  )
  )
  
better_plot
```
```{r}
ui <- fluidPage(
  sliderInput("year_slider", "Select Year", min = min(merged_data$year), max = max(merged_data$year), value = min(merged_data$year), step = 1),
  plotlyOutput("scatter_plot")
)


```

