---
title: "Are wealthier countries more digitized?"
author: "Dylan Jacobs, Jun Pu, and Ariel Lutati"
date: "2024-03-07"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Import packages
library(gapminder)
library(ggplot2)
library(plotly)
library(dplyr)
library(readxl)
library(leaflet)
library(sf)
library(RColorBrewer)
library(shiny)
```


# What question are we addressing?
- Are wealthier countries more digitalized? What is the relationship between status of economic development and financial digitalization?

# Why is it worth asking?
A digitalized financial system and electronic payments have the potential to contribute significantly to building a more resilient financial ecosystem in developing countries. This, in turn, can play a pivotal role in fostering economic development.

1. Increases account ownership through an enabling environment

2. Increases use of financial servieces: digital payments, borrowing & savings.

3. Building a more resilient financial ecosystem and improving financial well-being

# Data Description
```{r message=FALSE, warning=FALSE, include=FALSE}
# Import the two datasets: DatabankWide and df_gdpPC
df_findex <- read_excel("DatabankWide.xlsx")
df_gdp <- read.csv("df_gdpPC.csv")

# Select the variables of interest
df_findex_var <- df_findex[c("Country name", "Year", "Region", "Adult populaiton", "Income group", "Made or received a digital payment (% age 15+)", "Financial institution account (% age 15+)")]
df_gdp_var <- df_gdp[c("country", "year", "gdpPercap", "lifeExp")]

# Renmae variables
# Create a new dataframe df_rename
df_findex_new <- df_findex_var

# Rename "Made or received a digital payment (% age 15+)" to "payment_digital"
colnames(df_findex_new)[colnames(df_findex_new) == "Made or received a digital payment (% age 15+)"] <- "payment_digital"

# Rename "Financial institution account (% age 15+)" to "account_fin"
colnames(df_findex_new)[colnames(df_findex_new) == "Financial institution account (% age 15+)"] <- "account_fin"

# Rename "Country name" to "country"
colnames(df_findex_new)[colnames(df_findex_new) == "Country name"] <- "country"

# Rename "Adult populaiton" to "population"
colnames(df_findex_new)[colnames(df_findex_new) == "Adult populaiton"] <- "population"

# Rename "Income group" to "group_income"
colnames(df_findex_new)[colnames(df_findex_new) == "Income group"] <- "group_income"

# Rename "Year" to "year"
colnames(df_findex_new)[colnames(df_findex_new) == "Year"] <- "year"

# Rename "Region" to "region"
colnames(df_findex_new)[colnames(df_findex_new) == "Region"] <- "region"

# Remove N/A values
df_findex_new <- na.omit(df_findex_new)

# Keep observations from years 2014, 2017, and 2021
df_gdp_var_limit <- subset(df_gdp_var, year %in% c(2014, 2017, 2021))
df_findex_limit <- subset(df_findex_new, year %in% c(2014, 2017, 2021))

# Merge the two datasets by 'country' and 'year'
merged_data <- merge(df_findex_limit, df_gdp_var_limit, by = c("year", "country"))
```

## Which datasets are we using?

In this visualization project, we leverage two key datasets. Our primary dataset is [The Global Findex Database 2021](https://www.worldbank.org/en/publication/globalfindex), offering comprehensive country-level data on various financial indicators such as account ownership, payments, savings, credit, and more. Notably, the database reports indicators for the years 2021, 2017, 2014, and 2011. The second dataset we imported is GDP per Capita for the year 2017 downloaded from the [World Bank](https://data.worldbank.org/indicator/NY.GDP.PCAP.CD).

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
df_findex <- read_excel("DatabankWide.xlsx")
df_gdp <- read.csv("df_gdpPC.csv")
```

We merged the two datasets by "year" and "country".
```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
# Merge the two datasets by 'country' and 'year'
merged_data <- merge(df_findex_limit, df_gdp_var_limit, by = c("year", "country"))
```

## What are our variables of interest?
#### Made or received a digital payment (% age 15+)
The percentage of respondents who report using mobile money, a debit or credit card, or a mobile phone to make a payment from an account—or report using the internet to pay bills or to buy something online or in a store—in the past year.

#### Financial institution account (% age 15+)
The percentage of respondents above 15 years old who report having an account (by themselves or together with someone else) at a bank, credit union, microfinance institution, or post office that falls under prudential regulation by a government body.

#### Mobile money account (% age 15+)
The percentage of respondents who report personally using a mobile money service to make payments, buy things, or to send or receive money in the past year.

#### Has access to the internet (% age 15+)
The percentage of respondents who report having access to the internet in a country.

#### Received digital payments (% age 15+)
The percentage of respondents who report using a mobile money account, a debit or credit card, or a mobile phone to receive a payment into an account in the past year.

#### Owns a credit card, older (% age 25+)
The percentage of respondents who reports owning a credit card in the past year.

#### Income Group
Income group is a categorical variable that classifies countries into high income, upper-middle income, lower-middle income, and low income.

#### Region
Region is the continent the country is located in.

#### Economy
Economy is the name of the specific country.

### What are potential weaknesses of the data?
1) Limited years of observation: We're only looking at 2014, 2017, and 2021
2) Country-level data: aggregated data might mask significant variability within each country.

# Visualization 1 - Correlation between Digital Transactions and Ownership of Financial Account
The following shinyapp shows the relationship between percentage of population that has made/received a digital payment and percentage of population with an account at a financial institution. From the graph, we can observe that high income countries have a larger percentage with a financial account and a large percentage of involvement in digital payments.

```{r echo=FALSE, message=FALSE, warning=FALSE}
shiny::shinyAppFile("Visualization1_shinyapp.R")
```


# Visualization 2 - Comparison of Summary Statistics: Financial and Digital Access Across Income Groups
```{r echo=FALSE, message=FALSE, warning=FALSE}
findex2 <- df_findex %>%
  group_by(`Income group`) %>%
  mutate(avg_mobile_money = 100 * mean(`Mobile money account (% age 15+)`, na.rm = TRUE),
         avg_internet_access = 100 * mean(`Has access to the internet (% age 15+)`, na.rm = TRUE),
         avg_digital_payments = 100 * mean(`Received digital payments (% age 15+)`, na.rm = TRUE),
         avg_credit_card = 100 * mean(`Owns a credit card, older (% age 25+)`, na.rm = TRUE))

# Let's split up the countries by income group and examine various stats

findex3 <- findex2 %>%
  select(`Income group`, avg_mobile_money, avg_internet_access, avg_digital_payments, avg_credit_card)

findex4 <- findex3[c(1, 5, 9, 23),]


findex4 %>%
  plot_ly(x = ~`Income group`, y = ~avg_mobile_money, type = "bar", name = "Money Market Account") %>%
  add_trace(y = ~avg_internet_access, name = 'Internet Access') %>%
  add_trace(y = ~avg_digital_payments, name = 'Digital Payments') %>%
  add_trace(y = ~avg_credit_card, name = "Credit Card Ownership") %>%
  layout(yaxis = list(title = 'Percent'), barmode = 'group', xaxis = list(title = 'Income', categoryorder = "array",
                                                                  categoryarray = c("Low income", 
                                                                                    "Lower middle income", 
                                                                                    "Upper middle income",
                                                                                    "High income"))
  )
```


# Visualization 3 - Average Percentage of Digital Payments Across Regions and Income Groups
```{r echo=FALSE, message=FALSE, warning=FALSE}
shiny::shinyAppFile("Visualization3_shinyapp.R")
```


